---
title: "The Models"
author: "Ashley Bonner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse) #data library
library(deSolve) #ODE solver
library(patchwork) #figure arrangement
library(DiagrammeR) #flow charts
library(dplyr) #data wrangling

knitr::opts_chunk$set(echo = TRUE)
```

# The One-Pool model



Carbon in the soil can be represented as $C$ in grams per meter squared, and then we can describe the rate of change as the difference between the rate of inputs ($u(t)$, in grams per meter squared per year) and the rate of outputs (decay).
Decay is often considered to be proportional to the total carbon in the soil, so we cannot decay to a negative soil carbon content.
Thus, we see decay as a rate ($k$, per year), representing the inverse of turnover time ($\tau$, years).
At risk of over-simplification, turnover time can be considered the average amount of time organic carbon is expected to remain in the soil.

```{r Diagram}
grViz("
  digraph One_Pool_Model {
    rankdir=LR
    
    subgraph outside_the_system{ 
    node[ shape = oval,
          font = Baskerville,
          style=dashed]
    Inputs[label = 'Vegetation']; CarbonDioxide
    }
    
    
    Soil    -> CarbonDioxide [label = 'k']
    
    Inputs  ->  Soil [label = 'u(t)']

    
  subgraph cluster_SoilPools {
        label = 'Soil Pools'
          
        Soil[color=brown3, fontcolor=brown4]
    }
  }
")


```

We can describe the rate of change of soil carbon, then, with a simple ordinary differential equation:

$$\frac{dC}{dt} = u(t) - kC = u(t) - \frac{1}{\tau}C(t). $$
As a summary, we have one main driver of the system (reliant on dynamics outside of the soil):
- $u(t)$, or carbon inputs in units of mass concentration per time, specifically g-C meters$^{-2}$ year$^{-1}$, which can change over time. 

We also have one parameter (a value endemic to the soil itself):
 - $k$, a decay rate equal to the inverse of turnover time ($\tau$).

Below we set up this ODE model and, as an example, a constant function to represent inputs.
As inputs are driven by dynamics outside of our soil system, they can change over time, but for now we consider carbon inputs as averaged over time.

```{r OnePoolSetup}
#sources in the ODE model
source("R/OnePool_ODE.R")

#creates a constant function for spin up simulations, so we can graph the inputs against the model simulations.
constant.fn <- function(time, parms){
  return(parms$ave_inputs)
}


seasonal.fn <- function(time, parms){
  inputs <- (5/300)*sin((pi/6)*time+(3*pi/2))+(5/300) 
  return(inputs)
}

#params
Params_OnePool.ls <- list(ave_inputs = 260, #g-C m-2 yr-1
                                #u; [0, 2.5] kg-carbon m-2 yr-1
                          turnoverTime = 20
                                #\tau; [10, 50] yr
                          
)

Params_OnePool.ls$inputs.fn <- constant.fn

SimTimes <- seq(0, 100, by = 1/12) #one hundred years, monthly

```

## Equilibrium, Steady State

When ecosystems are mature, they are often considered to achieve a kind of equilibirum.
This is analogous to a concept in the mathematical study of ODEs: the steady state values.
Steady State values are the values of the state variables that cause the system to have a net zero rate of change; that is for our model, the ODE rate of change $\frac{dC}{dt} = 0$.
This occurs when the inputs and the outputs are balanced, when $u(t) = kC$, or:

$$\frac{dC}{dt} = 0 \implies C = u(t) \cdot \tau.$$

This is easily calculated and verified on a simulation run forward in time, where the total soil carbon concentration remains steady.

```{r SteadyState}

source("R/OnePool_SS.R")

Sim1_SS.df <- lsoda(y = OnePool_SS.fn(parms = Params_OnePool.ls),
                times = SimTimes,
                func = OnePool_ODE.fn,
                parms = Params_OnePool.ls) %>%
  as.data.frame() %>%
  mutate(Respiration_Rate = OnePool_ODE.fn(t = NA, 
                                           y = c(NA, Soil_Carbon_Concentration),
                                           parms = Params_OnePool.ls
                            )[[1]]['Cumulative_Respiration'],
          .by = everything()
    ) %>%
  mutate(Input_Rate = Params_OnePool.ls$inputs.fn(t = time,
                                                  parms = Params_OnePool.ls),
         .by = everything()
  )

#Verfication of Steady State:
ggplot(Sim1_SS.df %>% pivot_longer(cols = Cumulative_Respiration:Input_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  xlab("Time (years)") +
  ylab("Mass in kg C per m^2" )

```

One might notice that this gets far more complicated if the inputs are not considered constant.
In this case, $u(t) = kC(t)$ is not guaranteed to be true across any or all values of $t$ in a time interval.
Even in this case, we can consider a type of "quasi steady state" value across an average, and this will be important for some of the analysis we do below, but for now, let's double check that our forward run code works for a spin up simulations, and then consider a more involved model that aims to capture more of the dynamics which occur in soil.


## Example: Spin up simulation

Spinning up a model is often done on large Earth Systems Models to allow the model to replicate current state variable values.
That is, we assume for a basis to work from, that systems across the earth are mostly mature and are generally some kind of stable.
It is from here that models adjust drivers and parameters to create forecasts and predictions. 
Spin-ups start with constant drivers and state variables start at zero.

```{r}

Sim1.df <- lsoda(y = c("Cumulative_Respiration" = 0, "Soil_Carbon_Concentration" = 0),
                times = SimTimes,
                func = OnePool_ODE.fn,
                parms = Params_OnePool.ls) %>%
  as.data.frame() %>%
  mutate(Respiration_Rate = OnePool_ODE.fn(t = NA, 
                                           y = c(NA, Soil_Carbon_Concentration),
                                           parms = Params_OnePool.ls
                            )[[1]]['Cumulative_Respiration'],
          .by = everything()
    )%>%
  mutate(Input_Rate = Params_OnePool.ls$inputs.fn(t = time,
                                                  parms = Params_OnePool.ls),
         .by = everything()
  )

ggplot(Sim1.df %>% pivot_longer(cols = Cumulative_Respiration:Input_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  xlab("Time (years)") +
  ylab("Mass in kg C per m^2" )

```




# The Five-Pool Model: Century

A five-pool decay model featuring two litter pools and three soil pools with varying decay rates throughout, a similar structure to CENTURY [Parton 1987?], considered one of the seminal papers introducing this style of model.
The decay rates intrinsically represent several processes (mineral-organic associations, microbial uptake or facilitated transformations, moisture- or light-facilitated breakdown etc.), and the carbon that leaves one pool can either be transferred into a different pool or respired out of the system.
This is overwhelming the model structure of choice, albiet for different numbers of soil and litter pools, for global Earth system models (ESMs) [KTB 2012].


```{r Diagram, echo = FALSE}
grViz("
  digraph Five_Pool_Model {
    rankdir=LR
    
    subgraph outside_the_system{ 
    node[ shape = oval,
          font = Baskerville,
          style=dashed]
    Inputs[label = 'Vegetation']; CarbonDioxide
    }
    
    Inputs  -> Metabolic     [headport=w, color = goldenrod]; 
    Inputs  -> Structural    [headport=w, color = green3]; 
    Fast    -> CarbonDioxide [tailport=s, color = red3]; 
    Slow    -> CarbonDioxide [tailport=s, color=purple3]; 
    Passive -> CarbonDioxide [tailport=s, color=blue3]
    
    Metabolic  -> Fast          [color = goldenrod, tailport = e]
    Metabolic -> CarbonDioxide  [color = goldenrod, tailport = e]
    Structural -> Fast          [color = green3, tailport = e]
    Structural -> Slow          [color = green3, tailport = e]
    Structural -> CarbonDioxide [color = green3, tailport = e]
    
    subgraph Pools{
    
    
      subgraph cluster_LitterPools{
        label = 'Litter Pools'
        
        node[ shape = oval,
              font = Baskerville,
              style = solid]
          rank=same
          
        Metabolic[color=goldenrod, fontcolor=darkgoldenrod]; Structural[color=green3, fontcolor=green4]
              
      }
    
      subgraph cluster_SoilPools {
        label = 'Soil Pools'
    
        node[ shape = oval,
              font = Baskerville,
              style = solid]
        rank = same
          
        Fast[color=red3, fontcolor=red4]; Slow[color=purple3, fontcolor=purple4]; Passive[color=blue3, fontcolor=blue4]
        
        #Fluxes out of Fast
        Fast -> Slow[tailport=s, color = MediumVioletRed]
        Fast -> Passive[tailport=s, color = DarkViolet]
        
        #Fluxes out of Slow
        Fast -> Slow [dir=back, color = MediumVioletRed, headport = s]
        Slow -> Passive[tailport=s, color = SlateBlue3]
        
        #Fluxes out of Passive
        Fast -> Passive[dir=back, color=DarkViolet, headport = s]
        #Slow -> Passive[dir=back, color=SlateBlue3, headport = n]
      }
    }
    }
  }
")

```

Generally speaking, we can consider models where carbon can move from any pool for another; these arrows represent those fluxes present in CENTURY.

This diagram comes in the form of the matrix ordinary differential equation

$$\frac{d\vec{C}}{dt} = u \vec{b} - \boldsymbol{A}\boldsymbol{K}\vec{C} $$
where:
 - $\vec{C}$ is a vector representing the five pools (state variables) in the order from the fastest decay rate to the slowest (more on that later), 
 - $u$ is the inputs to the system (often estimated as a fraction of vegetation's net primary production),
 - $\vec{b}$ is a vector that divides up the carbon inputs into the soil system (in Century, this is all into the litter pools)
 - $\boldsymbol{K}$ is a diagonal matrix with the decay rates of the pools,
 - $\boldsymbol{A}$ is called the transfer matrix and can be of three different types [diagonal (for a model where the pools are independant and have no fluxes betwen them), lower-triangular (for a cascade model, where carbon flows from simpler forms to more complex forms due to the ecological principal of trophic cascades), or fully dense (for a model that allows carbon to move freely between the pools)]. In any of these three cases, the diagonal entries are all positive one, and the off-diagonal entries are nonpositive (negative or zero), and the columns will always sum to a positive value less than one. 

CENTURY as debuted in 1983 included the effects of temperature and moisture on decomposition, which was applied across all pools the same way, thus we omit it here as the effects of temperature will only serve to scale decomposition of the one-pool model the same as it does on this five-pool model.

```{r Parameters}

source("R/FivePool_ODE.R")

Params_FivePool.ls <- list(
                  ##Input-based parameters (such as allocation)
                  ave_inputs = 260,
                        #CENUTRY model uses 260 grams C per m^2 per year
                        #u; kg-carbon m-2 yr-1, 
                  # inputs.fn = constant.fn, ##Needs to be added in seperately
                    input_to_struc = 2/9,  
                    input_to_meta = 7/9,  
                        #C goes through litter pools first
                    input_to_fast = 0,  
                    input_to_slow = 0,  
                    input_to_passive = 0,  
                            
                  ##turnover times (years)
                     turnoverTime_meta = 3.5,  #this is CENTURY 1987, CENTURY 1983 has .1 to 1 year
                     turnoverTime_struc = 9.5, #adjusted out of initial range to match CENTURY model validation for total plant residue
                        # CENTURY: 1 to 5 years
                     turnoverTime_fast = 10, #this is CENTURY 1987, CENTURY 1983 has 10 years
                     turnoverTime_slow = 50, #adjusted, CENTURY 1987 has ~263, CENTURY 1983 has 50
                     turnoverTime_passive = 2500, #adjusted, CENTURY 1987 has ~7690, CENTURY 1983 has 1500 years
                     
                  ##Transfer Matrix values 
                       struc_to_fast = .7*.35, 
                       struc_to_slow = .3*.35, 
                          #60% lost to respiration, the remaining 40% is called Net Microbial Production 
                          #Adjusted to 75% lost to respiration to reach model validations percentages
                       meta_to_fast = .35,  
                          #60% lost to respiration, rest to active
                          #Adjusted to 75% lost to respiration to reach model validations percentages
                       fast_to_slow = .4,      
                          #60% lost to respiration, rest to slow
                       slow_to_fast = 0.92*.4,        
                       slow_to_passive = .08*.4,     
                          #60% lost to respiration, rest split between active and passive
                       passive_to_fast = .4,
                    #.4,
                          #60% lost to respiration, rest to slow
                    
                      #all the zero-flows
                        meta_to_struc = 0,
                        meta_to_slow = 0,
                        meta_to_passive = 0,
                        struc_to_meta = 0,
                        struc_to_passive = 0,
                        fast_to_passive = 0,
                        fast_to_meta = 0,
                        slow_to_meta = 0,
                        passive_to_meta = 0,
                        fast_to_struc = 0,
                        slow_to_struc = 0,     
                        passive_to_slow = 0,
                        passive_to_struc = 0
                       )


# # For other Markdowns:
# write_csv(as.data.frame(Params_FivePool.ls), file = "Data/FivePoolModel_Parameters.csv")
# 
# Params_FivePool.df <- read_csv(file.path("Data/FivePoolModel_Parameters.csv"), show_col_types = FALSE)
# Parms_FivePool.ls <- list()
# 
# for(i in 1:ncol(Params_FivePool.df)) {
#   Parms_FivePool.ls[[i]] <- as.numeric(Params_FivePool.df[,i])
# }
# names(Parms_FivePool.ls) <- colnames(Params_FivePool.df)


Params_FivePool.ls$inputs.fn <- constant.fn

```

##Example: Spin up simulation

```{r Simulation}
SimTimes <- seq(0, 8000, by = 1) 

Sim5.df <- lsoda(y = c(Cumulative_Respiration=0, 
                       Metabolic_Plant_Residue= 0,
                       Structural_Plant_Residue=0,
                       Fast_Soil = 0,
                       Slow_Soil = 0,
                       Passive_Soil = 0),
                      times = SimTimes,
                      func = FivePool_ODE.fn,
                      parms = Params_FivePool.ls) %>%
                  as.data.frame() %>%
                  mutate(C_total = Metabolic_Plant_Residue + Structural_Plant_Residue + 
                           Fast_Soil + Slow_Soil + Passive_Soil) %>%
                  mutate(Respiration_Rate = FivePool_ODE.fn(t = time, 
                                                            y = c(NA, 
                                                                  Metabolic_Plant_Residue, Fast_Soil, Structural_Plant_Residue, 
                                                                  Slow_Soil, Passive_Soil),
                                                            parms = Params_FivePool.ls
                                                            )[[1]]['Cumulative_Respiration'],
                          .by = everything())

ggplot(Sim5.df %>% 
         select(time, Cumulative_Respiration, C_total, Respiration_Rate) %>%
         pivot_longer(cols = c(Cumulative_Respiration, C_total, Respiration_Rate),
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  xlab("Time (years)") +
  ylab("Mass in kg C per m^2" )+
  scale_x_log10() 

temp <- Sim5.df %>% 
         select(time, Metabolic_Plant_Residue, Structural_Plant_Residue, Fast_Soil, Slow_Soil, Passive_Soil) %>%
         pivot_longer(cols = Metabolic_Plant_Residue:Passive_Soil,
                                  names_to = 'Pool',
                                  values_to = 'Mass')

temp$Pool <- factor(temp$Pool, levels = c("Metabolic_Plant_Residue", "Structural_Plant_Residue", "Fast_Soil", "Slow_Soil", "Passive_Soil"))

ggplot(temp) + 
  geom_line(aes(x=time, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') 

ggplot(temp) +
  geom_area(aes(x=time, y=Mass, fill = Pool)) +
  scale_fill_brewer(palette = "Spectral") 

ggplot(Sim5.df %>%
         mutate(Plant_Residue = Metabolic_Plant_Residue + Structural_Plant_Residue) %>%
         select(-c('Cumulative_Respiration', 'Respiration_Rate', 'Metabolic_Plant_Residue', 'Structural_Plant_Residue')) %>%
         pivot_longer(cols = Fast_Soil:Plant_Residue,
                                  names_to = 'Pool',
                                  values_to = 'Mass')) +
  geom_line(aes(x=time, y=Mass, linetype = Pool, color = Pool)) 


```

```{r SteadyState}
#Calls the Steady State function for the five pool model
source("R/FivePool_SS.R")

#Verification of Steady State:
FivePool_SteadyState <- as.data.frame(t(FivePool_SS.fn(parms=Params_FivePool.ls))) %>%
  mutate(Total_SOC = sum(FivePool_SS.fn(parms=Params_FivePool.ls)))

FivePool_SteadyState <- FivePool_SteadyState %>%
  mutate(Total_PR = Metabolic_Plant_Residue + Structural_Plant_Residue) %>%
  pivot_longer(cols = c(names(FivePool_SteadyState), "Total_PR"),
               names_to = "Pool",
               values_to = "gC_per_m2") %>%
  mutate(Percentages = 100*gC_per_m2/sum(FivePool_SS.fn(parms=Params_FivePool.ls))) 


FivePool_SteadyState
#Goal: Simulated: PR 16, Active 11, Slow 29, Passive 44
```


