---
title: "The Models"
author: "Ashley Bonner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse) #data library
library(deSolve) #ODE solver
library(patchwork) #figure arrangement
library(DiagrammeR) #flow charts
library(dplyr) #data wrangling

knitr::opts_chunk$set(echo = TRUE)
```

# The One-Pool model

Facts about it

$$\frac{dC}{dt} = u(t) - kC $$
or

$$\frac{dC}{dt} = u(t) - \frac{1}{\tau}C. $$

 - $u(t)$ is carbon inputs, which can change over time. This is in mass concentration per time, specifically kgC meters$^{-2}$ month$^{-1}$. 
 - $k$ is a decay rate equal to the inverse of turnover time.


```{r OnePoolSetup}

source("R/FOLD_OnePool_ODE.R")

#params
Params_OnePool.ls <- list(inputs = 0.2/12, 
                                #u; [0, 2.5] kg-carbon m-2 yr-1; split per month (to fit CENTURY parameters)
                              turnoverTime = 15*12
                                #\tau; [10, 50] yr, by month
                             ) 

SimTimes <- seq(0, 100*12, by = 1) #one hundred years, monthly

```

## Example: Spin up simulation

```{r}

Sim.df <- lsoda(y = c("Cumulative_Respiration" = 0, "Soil_Carbon_Concentration" = 0),
                times = SimTimes,
                func = OnePool_ODE.fn,
                parms = Params_OnePool.ls) %>%
  as.data.frame() %>%
  mutate(Respiration_Rate = OnePool_ODE.fn(t = NA, 
                                           y = c(NA, Soil_Carbon_Concentration),
                                           parms = Params_OnePool.ls
                            )[[1]]['Cumulative_Respiration'],
          .by = everything()
    )

ggplot(Sim.df %>% pivot_longer(cols = Cumulative_Respiration:Respiration_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time/12, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  xlab("Time (years)") +
  ylab("Mass in kg C per m^2" )

```



## Equilibrium, Steady State

$$\frac{dC}{dt} = 0 \implies C = u(t) \cdot \tau$$ 

```{r SteadyState}

SS_OnePool.fn <- function(parms){
    ans <- parms$inputs * parms$turnoverTime
  return(c(Cummulative_Respiration = 0, Soil_Carbon_Concentration = ans))
}

Sim_SS.df <- lsoda(y = SS_OnePool.fn(parms = Params_OnePool.ls),
                times = SimTimes,
                func = OnePool_ODE.fn,
                parms = Params_OnePool.ls) %>%
  as.data.frame() %>%
  mutate(Respiration_Rate = OnePool_ODE.fn(t = NA, 
                                           y = c(NA, Soil_Carbon_Concentration),
                                           parms = Params_OnePool.ls
                            )[[1]]['Cumulative_Respiration'],
          .by = everything()
    )


```




# The Five-Pool CENTURY Model

A five-pool decay model featuring two litter pools and three soil pools with varying decay rates throughout, a similar structure to CENTURY [Parton 1987?], considered one of the seminal papers introducing this style of model.
The decay rates intrinsically represent several processes (mineral-organic associations, microbial uptake or facilitated transformations, moisture- or light-facilitated breakdown etc.), and the carbon that leaves one pool can either be transferred into a different pool or respired out of the system.
This is overwhelming the model structure of choice, albiet for different numbers of soil and litter pools, for global Earth system models (ESMs) [KTB 2012 paper].


```{r Diagram, echo = FALSE}
grViz("
  digraph Three_Pool_Model {
    rankdir=LR
    
    subgraph outside_the_system{ 
    node[ shape = oval,
          font = Baskerville,
          style=dashed]
    Inputs[label = 'Vegetation']; CarbonDioxide
    }
    
    Inputs  -> Metabolic     [headport=w, color = goldenrod]; 
    Inputs  -> Structural    [headport=w, color = green3]; 
    Fast    -> CarbonDioxide [tailport=s, color = red3]; 
    Slow    -> CarbonDioxide [tailport=s, color=purple3]; 
    Passive -> CarbonDioxide [tailport=s, color=blue3]
    
    Metabolic  -> Fast          [color = goldenrod, tailport = e]
    Metabolic -> CarbonDioxide  [color = goldenrod, tailport = e]
    Structural -> Fast          [color = green3, tailport = e]
    Structural -> Slow          [color = green3, tailport = e]
    Structural -> CarbonDioxide [color = green3, tailport = e]
    
    subgraph Pools{
    
    
      subgraph cluster_LitterPools{
        label = 'Litter Pools'
        
        node[ shape = oval,
              font = Baskerville,
              style = solid]
          rank=same
          
        Metabolic[color=goldenrod, fontcolor=darkgoldenrod]; Structural[color=green3, fontcolor=green4]
              
      }
    
      subgraph cluster_SoilPools {
        label = 'Soil Pools'
    
        node[ shape = oval,
              font = Baskerville,
              style = solid]
        rank = same
          
        Fast[color=red3, fontcolor=red4]; Slow[color=purple3, fontcolor=purple4]; Passive[color=blue3, fontcolor=blue4]
        
        #Fluxes out of Fast
        Fast -> Slow[tailport=s, color = MediumVioletRed]
        Fast -> Passive[tailport=s, color = DarkViolet]
        
        #Fluxes out of Slow
        Fast -> Slow [dir=back, color = MediumVioletRed, headport = s]
        Slow -> Passive[tailport=s, color = SlateBlue3]
        
        #Fluxes out of Passive
        Fast -> Passive[dir=back, color=DarkViolet, headport = s]
        #Slow -> Passive[dir=back, color=SlateBlue3, headport = n]
      }
    }
    }
  }
")

```

Generally speaking, we can consider models where carbon can move from any pool for another; these arrows represent those fluxes present in CENTURY.



