---
title: "Aggregate Decay Rate"
author: "Ashley Bonner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# An Aggregate Decay Rate

## Introduction

Through some mathematical theory, we can actually show that the product of the transfer and decay matrix ($\boldsymbol{KA}$) is an invertible matrix. 
This means that the equation can be solved for steady state, and the form this takes is:
$$ \vec{C}(t) = u \left(\boldsymbol{KA}\right)^{-1} \vec{b} $$
And if we sum up the pools, this actually gives us a single value for the steady state of total soil carbon in the model.
$$ \boldsymbol{C} = u \sum \left( \left( \boldsymbol{KA}\right)^{-1} \vec{b} \right) $$
Recall from the one pool model that the steady state was the carbon inputs ($u$) times the decay rate - allowing us to ask the question: can we consider $\sum \left( \left( \boldsymbol{KA}\right)^{-1} \vec{b} \right)$ to be a proxy decay rate for a more complicated first-oder linear decay model, thus allowing us to use a simpler one-pool model implimentation instead?

We start by calculating what this aggregrate decay rate would be for the Century parameters. 
One of the cool bits about this method is that the size of the vector and matrix values do not matter - as long as you ensure that the decay rates on the matrix are in descening order (longer turnover times last).
This reduced complexity one pool model can be used in place of a first-order linear decay model with any number of pools.

```{r ProxyRate}
#Pull in the aggregate decay rate function
source("R/AggregateDecayRate.R")

#Create a Parameter list for the One-pool model using the decay rate and the same inputs
Params_Aggregate.ls <- list(ave_inputs = 260, #g-C m-2 yr-1, according to CENTURY model validation
                            turnoverTime = Proxy_decayrate.fn(parms = Params_FivePool.ls)
                            )

#Define a Constant Function for initial test runs
constant.fn <- function(time, parms){
  return(parms$ave_inputs)
}

#Add needed input function
Params_Aggregate.ls$inputs.fn <- constant.fn

#Simulation times for a spin up example of the aggrgate
SimTimes <- seq(0,300, by=1)

#Simulation for a spin-up simulation with the one-pool model
SpinUp_Proxy_sim.df <- lsoda(y = c("Cumulative_Respiration" = 0, 
                                    "Soil_Carbon_Concentration" = 0),
                                    times = SimTimes,
                                    func = OnePool_ODE.fn,
                                    parms = Params_Aggregate.ls) %>%
                          as.data.frame() %>%
                  mutate(RespirationRate = OnePool_ODE.fn(t = NA, 
                                                          y = c(NA, Soil_Carbon_Concentration),
                                                          parms = Params_Aggregate.ls)[[1]]['Cumulative_Respiration'],
                          .by = everything()
                  )


ggplot(SpinUp_Proxy_sim.df) + 
  geom_line(aes(x=time, y=Soil_Carbon_Concentration)) +
  #This additional lsoda run verifies that the steady state found for the Five pool model remains the steady state under our aggregate decay rate
  geom_line(data = lsoda(y = c(Cumulative_Respiration=0, 
                                Soil_Carbon_Concentration = sum(FivePool_SS.fn(parms=Params_FivePool.ls))),
                          times = SimTimes,
                          func = OnePool_ODE.fn,
                          parms = Params_Aggregate.ls) %>%
                            as.data.frame() , 
            aes(x=time, y=Soil_Carbon_Concentration), linetype="dotted") 

ggplot(SpinUp_Proxy_sim.df) + 
  geom_line(aes(x=time, y=RespirationRate)) +
  geom_line(data = lsoda(y = c(Cumulative_Respiration=0,
                                Soil_Carbon_Concentration = sum(FivePool_SS.fn(parms=Params_FivePool.ls))),
                          times = SimTimes,
                          func = OnePool_ODE.fn,
                          parms = Params_Aggregate.ls) %>%
                            as.data.frame() %>%
                            mutate(RespirationRate = OnePool_ODE.fn(t = NA, 
                                                          y = c(NA, Soil_Carbon_Concentration),
                                                          parms = Params_Aggregate.ls)[[1]]['Cumulative_Respiration'],
                                  .by = everything()
                                ),
            aes(x=time, y=RespirationRate), linetype="dotted")

```

# How does it compare?

Does a single pool model structure with first order linear decay actually stand up against a more robust multi-pool structure?
In the Scenario Markdowns, we test this aggregate decay rate for a one pool model against a five pool Century model and compare the outputs.

