---
title: "FOLD figure generation"
author: "Kathe Todd-Brown, Ashley Bonner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
#useful functions in here

library(tidyverse) #data library
library(deSolve) #ODE solver
library(DiagrammeR) #flow charts

knitr::opts_chunk$set(echo = TRUE)
```

# One pool model

The rate of change of soil carbon stocks is the balance between the inputs and the outputs because of conservation of mass -- carbon is neither created nor distroyed.
In this simple one pool model we assume a constant input rate ($u$; [mass per area per time]).
The outputs are assumed to be proportional to the stock that is there.
There are other functional forms that could also use but this is the simplest (and also happens to fit incubation experiments).

$\frac{dC}{dt} = u - kC$

where 

  + $C$ is the carbon stock [mass per area]. This is often on the order of 1 to 150 kg-organic-carbon m^-2^ for the first meter of soil (HWSD2)
  + $u$ is the inputs into the soil [mass per area per time]. Between root exodase, root turnover, and surface litter, inputs are difficult to quantify directly. However net primary productivity is often used as a proxy for soil inputs and varies between 0 to 2.5 kg-organic-carbon m^-2^ yr^-1^ with the geographic mode occurring around 0.2 kg-C m^-2^ yr^-1^ (MODIS - MOD17A3)
  + $k$ is the decay rate is the fraction of carbon leaving the soil carbon pool per time. This is often framed as it's inverse, turnover time ($k=\frac{1}{\tau}$) which has many interpretation but for this model is the average length of time a carbon atom will spend in the soil. The bulk turnover time in Earth system models is on the order of 10 to 50 years (CMIP5 and CMIP6).
  
This turnover time is so commonly used that we will re-frame this equation:

$\frac{dC}{dt} = u - \frac{C}{\tau}$

From this model we can calculate the steady state value where the rate of change is zero.
This describes the carbon stock that the system is trending towards as it matures.

$C_{ss} = \frac{u}{k} = u \tau$

This also gives us a useful check on our ranges if we assume that soil carbon stocks are generally close to steady state. 
$u\tau = [0, 0.2, 2.5] [10, 15, 50] = [0, 3, 125] $
This is reassuring! 
It's very very close to the stated ranges for the soil carbon stocks of $[1,150]$.

We can also look at the inputs and carbon stocks to infer the turnover times.
$\frac{C}{u} = \frac{[1, 150]}{[[0.1, 2.5]]} = [0.4, 1500]$
which is a MUCH larger range of turnover times.
So this would be consistent with the observations but maybe not constraining of the parameters.

So let's code this up!

```{r}
 #u; [0, 2.5] kg-carbon m-2 yr-1;
parameter.ls <- list(inputs = 0.2, 
                     #\tau; [10, 50] yr
                     turnoverTime = 15) 

#rate change function
dCdt_1pool <- function(t, y, parms){
  ans <- parms$inputs - y / parms$turnoverTime
  return(list(ans))
}

#steady state function
C_ss_1pool <- function(parms){
  ans <- parms$inputs * parms$turnoverTime
  return(c(C = ans))
}

# Numerically solving this is a little bit overkill since 
# ...the differential equation is solvable but for generalizations here is the numerical
# ...solution from the function lsoda in the deSolve package
sim_1pool <- lsoda(y = C_ss_1pool(parms = parameter.ls)*0.8,
                times = 0:100,
                func = dCdt_1pool,
                parms = parameter.ls) %>%
  as.data.frame() 

ggplot(sim_1pool) +
  geom_line(aes(x=time, y=C))
```
# Three pool models

The rest of the documentation here is sparse but here is the general gist of the 'Three pool models' section.
Based on soil lab incubation and the isotopic age of the carbon, there is evidence that a one pool model can't match the respiration curves over multi-year incubation and observed soil carbon ages.
This has led to a three pool model where the soil carbon stock is split into three pools with different turnover times.
The turnover times for each pool are generally an order of magintude different from each other.

We then present the following set of equations (moving back to $k_i = 1/\tau_i$ for notation simplicity):

$\frac{C_f}{dt} = b_f u - k_fC_f + f_{sf}k_sC_s + f_{pf}k_pC_p $
$\frac{C_s}{dt} = b_s u + f_{fs}k_fC_f - k_sC_s + f_{ps}k_pC_p $
$\frac{C_p}{dt} = (1-b_f-b_s) u + f_{fp}k_fC_f + f_{sp}k_sC_s - k_pC_p $

where
  + $b_i$ is the allocation of the inputs to that pool $i$
  + $f_ij$ is the fraction of carbon from pool $i$ going to pool $j$ of the exiting pool $i$ carbon
  + $k_f > k_s > k_p$ the fast carbon pool ($C_f$) has a higher decay rate then the slow carbon pool ($C_s$) that in turn has a higher decay rate then the passive carbon pool ($C_p$). Often $\tau_f$ is on the order of years, $\tau_s$ is on the order of decades, $\tau_p$ is on the order of centuries to millennium.

For those of you who remember matrix multiplication this can be boiled down to the following.
[[Note I flip the dang matricies all the time so this might be wrong, need to double check the ordering of terms here. -KTB 2023 May 5]]

$$\frac{d}{dt}\left[\begin{array}{c} 
C_f \\ C_s \\ C_p
\end{array}\right] = \left[\begin{array}{c} 
b_f \\ b_s \\ 1-b_f-b_s
\end{array}\right] u - \left[\begin{array}{ccc} 
-k_f & f_{sf}k_s & f_{pf}k_p \\ f_{fs}k_f & - k_s & f_{ps}k_p \\ f_{fp}k_f & f_{sp}k_s & -k_p
\end{array}\right] \left[\begin{array}{c} 
C_f \\ C_s \\ C_p
\end{array}\right]$$

$$\frac{d}{dt}\left[\begin{array}{c} 
C_f \\ C_s \\ C_p
\end{array}\right] = \left[\begin{array}{c} 
b_f \\ b_s \\ 1-b_f-b_s
\end{array}\right] u - \left[\begin{array}{ccc} 
k_f & 0 & 0 \\ 0 & k_s & 0 \\0 & 0 & k_p
\end{array}\right] 
\left[\begin{array}{ccc} 
-1 & f_{sf} & f_{pf} \\ f_{fs} & - 1 & f_{sf} \\ f_{fp} & f_{sp} & -1
\end{array}\right] 
\left[\begin{array}{c} 
C_f \\ C_s \\ C_p
\end{array}\right] $$

$$\frac{d\vec{C}}{dt} = \vec{b} u - \textbf{K}\textbf{A}\vec{C}$$


In the sections below we go through three cases with a one-pool proxy estimate in each:

  + Independent - there is no transfer between pools and only the main diagonal of $\textbf{A}$ is non-zero or $-1$.
  + Cascade - carbon only moves from faster to slower pools leading to a lower triangular matrix $\textbf{A}$.
  + Full cascade - carbon is exchanged between all the pools leading to a fully dense matrix $\textbf{A}$.

## Independent pools


```{r}
 #u; [0.1, 4] kg-carbon m-2 yr-1;
#parameters.ls <- list(inputs = 1, turnoverTime = 15)

parameter.ls <- c( parameter.ls, 
                   list( 
                     input_to_fast = 0.55,
                     input_to_slow = 0.425,
                     #fast turnover \tau; [0, 10] yr
                     turnoverTime_fast = 3,
                     #slow turnover \tau; [10, 100] yr
                     turnoverTime_slow = 40,
                     #passive turnover \tau: [100, Inf] yr
                     turnoverTime_passive = 200) )


C_ss_3independent <- function(parms){
  C_fast <- parms$inputs * 
    parms$input_to_fast * 
    parms$turnoverTime_fast
  
  C_slow <- parms$inputs * 
    parms$input_to_slow * 
    parms$turnoverTime_slow
  
  C_passive <- parms$inputs * 
    (1-parms$input_to_fast - parms$input_to_slow) * 
    parms$turnoverTime_passive
  
  return(c(C_fast = C_fast,
           C_slow = C_slow,
           C_passive = C_passive))
}

#C_ss_3independent(parms = parameter.ls)

dCdt_3independent <- function(t, y, parms){
  dC_fast <- parms$inputs *parms$input_to_fast -
    y[1]/parms$turnoverTime_fast
  
  dC_slow <- parms$inputs * parms$input_to_slow - 
    y[2]/parms$turnoverTime_slow
  
  dC_passive <- parms$inputs * 
    (1-parms$input_to_fast - parms$input_to_slow) - 
    y[3]/parms$turnoverTime_passive
  
  return(list(c(C_fast = dC_fast,
           C_slow = dC_slow,
           C_passive = dC_passive)))
}


#dCdt_3independent(t=0, y = c(1,3,5), parms = parameter.ls)


sim_3independent <- lsoda(y = C_ss_3independent(parms = parameter.ls)*0.2,
                times = 0:1000,
                func = dCdt_3independent,
                parms = parameter.ls) %>%
  as.data.frame() %>%
  pivot_longer(cols = C_fast:C_passive,
               names_to = 'pool',
               values_to = 'stock')%>%
  mutate(pool = factor(pool, levels = c('C_fast', 'C_slow', 'C_passive')),
         model = '3independent')

ggplot(sim_3independent) +
  geom_line(aes(x=time, y=stock)) +
  facet_wrap(~pool)
```

```{r Ind_OnePool_Proxy}
Ind_SS <- C_ss_3independent(parms=parameter.ls)

sim_Ind_OnePoolProxy <- sim_3independent %>%
  pivot_wider(names_from="pool", values_from="stock") %>%
  mutate(C_Total = C_fast + C_slow +C_passive) %>%
  mutate(One_pool_proxy = Ind_SS[1] + C_slow + Ind_SS[3]*0.2) %>%
  select(time, C_Total, One_pool_proxy) 

ggplot(sim_Ind_OnePoolProxy) +
  geom_line(aes(x=time, y=C_Total, colour="blue")) +
  geom_line(aes(x=time, y=One_pool_proxy, colour="red")) +
  geom_vline(xintercept = 10) +
  geom_vline(xintercept = 100) +
  geom_hline(yintercept = Ind_SS[1]+Ind_SS[2]+Ind_SS[3], color="red", alpha=0.2, linetype=2) +
  geom_hline(yintercept = Ind_SS[1]+Ind_SS[2]+Ind_SS[3]*0.2, color="blue", alpha=0.2, linetype=2) 

plot_sim_Ind_OnePoolProxy <- sim_Ind_OnePoolProxy %>%
  filter(time > 10) %>%
  filter(time < 100)

ggplot(plot_sim_Ind_OnePoolProxy) +
  geom_line(aes(x=time, y=C_Total, colour="blue")) +
  geom_line(aes(x=time, y=One_pool_proxy, colour="red")) 
```
z
## Cascade Three-pool model

This model has a transfer matrix that is lower triangular. In the system, this means that as carbon leaves faster-decaying pools, some is respired as CO$_2$ and some becomes a part of more complex, thus slower decaying, forms of carbon in the soil (this process is known as the trophic cascade). 

```{r Three_Pool_Cascade}
## Independent pool parameters
parameter_independent <- c(parameter.ls, 
                  list(fast_to_slow = 0,
                     fast_to_passive = 0,
                     slow_to_passive = 0)) 

##3pool_cascade
parameter_cascade <- c(parameter.ls, 
                  list(fast_to_slow = 0.2,
                     fast_to_passive = 0.1,
                     slow_to_passive = 0.3)) 


C_ss_3cascade <- function(parms){
  C_fast <- parms$inputs * 
    parms$input_to_fast * 
    parms$turnoverTime_fast
  
  C_slow <- (parms$inputs * parms$input_to_slow +
               C_fast / parms$turnoverTime_fast * parms$fast_to_slow) *
    parms$turnoverTime_slow
  
  C_passive <- (parms$inputs * (1-parms$input_to_fast - parms$input_to_slow) +
      C_fast / parms$turnoverTime_fast * parms$fast_to_passive +
      C_slow / parms$turnoverTime_slow * parms$slow_to_passive )* 
    parms$turnoverTime_passive
  
  return(c(C_fast = C_fast,
           C_slow = C_slow,
           C_passive = C_passive))
}

if(all(C_ss_3independent(parms = parameter_cascade) == 
       C_ss_3cascade(parms = parameter_independent))){
  print('cross check with C_ss independent and cascade works')
}else{
  print('problem with C_ss cascade steady state check')
}

C_init <- C_ss_3cascade(parms = parameter_independent)

dCdt_3cascade <- function(t, y, parms){
  dC_fast <- (parms$inputs *parms$input_to_fast) -
    y[1]/parms$turnoverTime_fast
  
  dC_slow <- (parms$inputs * parms$input_to_slow +
                y[1] / parms$turnoverTime_fast * parms$fast_to_slow)- 
    y[2]/parms$turnoverTime_slow
  
  dC_passive <- (parms$inputs * (1-parms$input_to_fast - parms$input_to_slow) +
                   y[1] / parms$turnoverTime_fast * parms$fast_to_passive +
                   y[2] / parms$turnoverTime_slow * parms$slow_to_passive ) - 
    y[3]/parms$turnoverTime_passive
  
  return(list(c(C_fast = dC_fast,
           C_slow = dC_slow,
           C_passive = dC_passive)))
}

if(all(dCdt_3independent(t=0, y = C_init,parms = parameter.ls)[[1]] ==
       dCdt_3cascade(t=0, y = C_init,parms = parameter_independent)[[1]])){
  print('cross check with dCdt independent and cascade works')
}else{
  print('problem with dCdt cascade steady state check')
}

#dCdt_3cascade(t=0, y = C_init, parms = parameter.ls)


sim_3cascade <- lsoda(y = C_ss_3cascade(parms = parameter_cascade) *0.8,
                      times = 0:1000,
                      func = dCdt_3cascade,
                      parms = parameter_cascade) %>%
  as.data.frame() %>%
  pivot_longer(cols = C_fast:C_passive,
               names_to = 'pool',
               values_to = 'stock')%>%
  mutate(pool = factor(pool, levels = c('C_fast', 'C_slow', 'C_passive')),
         model = '3cascade')

ggplot(sim_3cascade) +
 geom_line(aes(x=time, y=stock, color = model)) +
  facet_wrap(~ pool)
```

###Cascade to Inpendent method
The cascade model
$$\frac{d\overrightarrow{C}}{dt} = u\overrightarrow{b} - \boldsymbol{K}\boldsymbol{A}\overrightarrow{C}$$
with $\boldsymbol{A}$ being lower triangular has a reformation that creates an independent model with the same totals. $\boldsymbol{KA}$ has an eigen decomposition $\boldsymbol{KA} = \boldsymbol{P \Lambda P^{-1}}$, where $\boldsymbol{P}$ is invertible, composed of the eigenvectors of $\boldsymbol{KA}$ and $\boldsymbol{\Lambda}}$ is a diagonal matrix with the corresponding eigenvalues. 
We let $\vec{\tilde{C}} = \boldsymbol{P^{-1}}\vec{C}$, then by replacing $\boldsymbol{KA}$ with its decomposition and then left-multiplying by $\boldsymbol{P}^{-1}$, we get that our new formulation is:

$$ \frac{d \overrightarrow{\tilde{C}}}{dt} = u\boldsymbol{P}^{-1}\overrightarrow{b} - \boldsymbol{\Lambda}\overrightarrow{\tilde{C}} $$

```{r Cascade2Ind}

#A "new" three-pool Independent model based on the prior Cascade model

dCdt_Cascade2Ind <- function(t, y, parms){
  
  decay_matrix <- diag(x= 1/c(parms$turnoverTime_fast,
                             parms$turnoverTime_slow,
                             parms$turnoverTime_passive))
  transfer_matrix <- matrix(c(1,  0, 0,
                              -parms$fast_to_slow, 1, 0,
                              -parms$fast_to_passive, -parms$slow_to_passive,  1),
                            nrow = 3,
                            byrow=TRUE)
  
     ##P^(-1) in proof
  Redistribution_matrix <- solve(eigen(decay_matrix %*% transfer_matrix)$vectors)   ##P^(-1) in proof
  
      ##Calculates new pool initial values
  new_pools <- as.list(Redistribution_matrix %*% y )                                ##Calculates new pool initial values
  names(new_pools)<-c("C_new_fast", "C_new_slow", "C_new_passive")
  
     ##Calculates new input fraction
  new_inputs <- Redistribution_matrix %*% c(parms$input_to_fast, parms$input_to_slow, 1-parms$input_to_fast-parms$input_to_slow) 
  names(new_inputs)<-c("new_input_to_fast", "new_input_to_slow", "new_input_to_passive")    
  
  
  new_decay_matrix <- diag(eigen(decay_matrix %*% transfer_matrix)$values)
  
  return(c(new_pools$C_new_fast, new_pools$C_new_slow, new_pools$C_new_passive))
  
}

dCdt_Cascade2Ind(y = C_ss_3cascade(parms = parameter_cascade)*0.8, parms=parameter_cascade)
  
  
  
#   dC_new_fast <- #(parms$inputs *parms$input_to_fast) -
#     #y[1]/parms$turnoverTime_fast
#   
#   dC_new_slow <- #(parms$inputs * parms$input_to_slow +
#                # y[1] / parms$turnoverTime_fast * parms$fast_to_slow)- 
#    # y[2]/parms$turnoverTime_slow
#   
#   dC_new_passive <- #(parms$inputs * (1-parms$input_to_fast - parms$input_to_slow) +
#                    #y[1] / parms$turnoverTime_fast * parms$fast_to_passive +
#                    #y[2] / parms$turnoverTime_slow * parms$slow_to_passive ) - 
#     #y[3]/parms$turnoverTime_passive
#   
#   return(list(c(C_fast = dC_new_fast,
#            C_slow = dC_new_slow,
#            C_passive = dC_new_passive)))
# }
```


```{r Cascade_OnePool_Proxy}
# decay_matrix <- diag(c(1/parameter.ls$turnoverTime_fast, 1/parameter.ls$turnoverTime_slow, 1/parameter.ls$turnoverTime_passive))
# 
# allocation_matrix <- matrix(c(1, 0, 0,
#                               -parameter.ls$fast_to_slow, 1, 0, 
#                               -parameter.ls$fast_to_passive, -parameter.ls$slow_to_passive, 1), 
#                       ncol = 3, byrow=TRUE)
# 
# comp_matrix <- decay_matrix %*% allocation_matrix
# 
# adjustment_matrix <- eigen(comp_matrix)$vectors
# 
# new_decay_rates <- diag(eigen(comp_matrix)$values)
# 
# #comp_matrix == adjustment_matrix %*% new_decay_rates %*% solve(adjustment_matrix)
# 
# 
# ##from above
# 
# 
# dCdt_3cascade <- function(t, y, parms){
#   
#   decay_allo_matrix <- matrix(data= c( parms$))
#   
#   
#   dC_fast <- (parms$inputs *parms$input_to_fast) -
#     y[1]/parms$turnoverTime_fast
#   
#   dC_slow <- (parms$inputs * parms$input_to_slow +
#                 y[1] / parms$turnoverTime_fast * parms$fast_to_slow)- 
#     y[2]/parms$turnoverTime_slow
#   
#   dC_passive <- (parms$inputs * (1-parms$input_to_fast - parms$input_to_slow) +
#                    y[1] / parms$turnoverTime_fast * parms$fast_to_passive +
#                    y[2] / parms$turnoverTime_slow * parms$slow_to_passive ) - 
#     y[3]/parms$turnoverTime_passive
#   
#   return(list(c(C_fast = dC_fast,
#            C_slow = dC_slow,
#            C_passive = dC_passive)))
# }
# 
# 
# 
# 
# # sim_3cascadereduction <- #*need to code this up*
# # 
# # sim_Cas_OnePoolProxy <- sim_3cascade_reduction%>%
# #   pivot_wider(names_from="pool", values_from="stock") %>%
# #   mutate(C_Total = C_fast + C_slow +C_passive) %>%
# #   mutate(One_pool_proxy = Ind_SS[1] + C_slow + Ind_SS[3]*0.8) %>%
# #   select(time, C_Total, One_pool_proxy) #%>%
# # #  filter(time > 9) %>%
# # #  filter(time <100)
# 

```

## Feedback model

```{r}

##3pool_feedback

#for generation of the intial states:
parameter_independent <- c(parameter_independent,
                  list(passive_to_fast = 0,
                       passive_to_slow = 0,
                       slow_to_fast = 0
                       ))

#Notes on the input fraction calculations (gleaned from Parton's CENTURY model)
  #Parton uses MONTHLY time steps
  #Century breaks up plant residue into metabolic C and Structural C based on the Lignin to Nitrogen ratio.
    #We will be using a 7/1 Lignin to Nitrogen ratio, thus 12.5% of plant litter is in the form of metabolic C and 87.5% is a structural C.
    #Century then sends 45% of metabolic C to the fast pool (the rest decays before entering the soil pool). - including this fraction breaks mass balance so it is omitted
 #Century then sends a portion of structural C to the fast and slow pools based on the lignin fraction A.
    ### !!!! because I have no idea what the lignin fraction even is, I'm using the placeholder value of 50%.
    # 70% of A* Structural C goes to the Slow pool (rest decays before reaching soil) - including this fraction breaks mass balance so it is omitted
    # 45% (?) of (1-A)*Structural C goes to fast (rest decays before reaching soil) - including this fraction breaks mass balance so it is omitted

century_parameter.ls <- list(inputs = 0.2/12, #monthly; divided evenly between the months
                            input_to_fast = ((.125)+(.875*.5)),  
                            input_to_slow = .875*.5,  
                                                   #note: Century includes no direct inputs to the passive pool.
                     #fast turnover according to Century \tau; [2, 4] yr
                     turnoverTime_fast = 18,    #1.5y  
                     #slow turnover according to Century \tau; [20, 50] yr
                     turnoverTime_slow = 300,       #25y
                     #passive turnover according to Century \tau: [800, 1200] yr
                     turnoverTime_passive = 12000,  #1000y
                     
                       fast_to_slow = 0.486,       # This is determined in CENTURY based on a soil vs. silt & clay fraction
                                                        # if soil to silt+clay fraction is zero, then = .15-.004, if fraction is 1, then =.83-.004
                                                        #currently using the average between the two: .486
                       fast_to_passive = 0.004,    #
                       slow_to_fast = 0.42,        #
                       slow_to_passive = 0.03,     #
                       passive_to_fast = 0.45,     #
                       passive_to_slow = 0         # CENTURY does not include this flow
                       )



C_ss_3feedback <- function(parms){
  allocation_vector <- matrix(c(parms$input_to_fast, 
                                parms$input_to_slow, 
                                1 - parms$input_to_fast - parms$input_to_slow),
                              nrow = 3)
  decay_matrix <- diag(x= 1/c(parms$turnoverTime_fast,
                             parms$turnoverTime_slow,
                             parms$turnoverTime_passive))
  transfer_matrix <- matrix(c(1,  -parms$slow_to_fast, -parms$passive_to_fast,
                              -parms$fast_to_slow, 1, -parms$passive_to_slow,
                              -parms$fast_to_passive, -parms$slow_to_passive,  1),
                            nrow = 3,
                            byrow=TRUE)
  #0 = inputs * allocation - decay_matrix %*% transfer_matrix %*% y
  #decay_matrix %*% transfer_matrix %*% y = inputs * allocation
  
  ans <- solve(decay_matrix  %*% transfer_matrix ,parms$inputs * allocation_vector)
  
  return(c(C_fast = ans[1],
           C_slow = ans[2],
           C_passive = ans[3]))
}


dCdt_3feedback <- function(t, y, parms){
  pools <- matrix(y, nrow=3)
 allocation_matrix <- matrix(c(parms$input_to_fast, 
                                parms$input_to_slow, 
                                1 - parms$input_to_fast - parms$input_to_slow),
                              nrow = 3)
  decay_matrix <- diag(x= 1/c(parms$turnoverTime_fast,
                             parms$turnoverTime_slow,
                             parms$turnoverTime_passive))
  transfer_matrix <- matrix(c(1,  -parms$slow_to_fast, -parms$passive_to_fast,
                              -parms$fast_to_slow, 1, -parms$passive_to_slow,
                              -parms$fast_to_passive, -parms$slow_to_passive,  1),
                            nrow = 3,
                            byrow=TRUE)
  #dcdt = inputs * allocation + transfer_matrix %*% decay_matrix %*% y
  #decay_matrix %*% transfer_matrix %*% y = - inputs * allocation
  
  ans <- parms$inputs * allocation_matrix - decay_matrix %*% transfer_matrix %*% y
  
  return(list(c(C_fast = ans[1],
           C_slow = ans[2],
           C_passive = ans[3])))
}


sim_3feedback <- lsoda(y = C_ss_3feedback(parms = century_parameter.ls)*0.6,
                      times = 0:(12*1000),
                      func = dCdt_3feedback,
                      parms = century_parameter.ls) %>%
  as.data.frame() %>%
  pivot_longer(cols = C_fast:C_passive,
               names_to = 'pool',
               values_to = 'stock')%>%
  mutate(pool = factor(pool, levels = c('C_fast', 'C_slow', 'C_passive')),
         model = '3feedback')

# #Full Feedback graph by pools:
# ggplot(sim_3feedback #%>%
#          #bind_rows(sim_3cascade, sim_3independent)
#        ) +
#   geom_line(aes(x=time, y=stock, color = model)) +
#   facet_wrap(~ pool)

```

```{r One_Pool_Proxy_Decay_Rate}
##One Pool Proxy Method:

#Proxy Decay Rate Calculation
Proxy_decayrate <- function(parms){
  decay_matrix <- diag(x= 1/c(parms$turnoverTime_fast,
                             parms$turnoverTime_slow,
                             parms$turnoverTime_passive))
  
  transfer_matrix <- matrix(c(1,  -parms$slow_to_fast, -parms$passive_to_fast,
                              -parms$fast_to_slow, 1, -parms$passive_to_slow,
                              -parms$fast_to_passive, -parms$slow_to_passive,  1),
                            nrow = 3,
                            byrow=TRUE)
  
   allocation_vector <- matrix(c(parms$input_to_fast, 
                                parms$input_to_slow, 
                                1 - parms$input_to_fast - parms$input_to_slow),
                              nrow = 3)
  
  ans <- sum(solve(decay_matrix %*% transfer_matrix) %*% allocation_vector)
  return(ans)
}  


##Adds the proxy rate to the main parameter list
century_parameter.ls$proxy_decay_rate <- Proxy_decayrate(parms=century_parameter.ls)
  


#One-Pool rate of change function
dCdt_FeedbackProxy <- function(t, y, parms){
  ans <- parms$inputs - y / parms$proxy_decay_rate
  return(list(ans))
}


#Initial Value for the One Pool Proxy, simulation, with name for lsoda output
C_init <- sum(C_ss_3feedback(parms = century_parameter.ls)) %>% setNames("C")

#Simulation with the one-pool model
sim_Feedback_OnePool <- lsoda(y = C_init*0.6,
                times = 0:(12*1000),
                func = dCdt_FeedbackProxy,
                parms = century_parameter.ls) %>%
  as.data.frame()


#Calculating total C in the Feedback run
sim_3feedback <- sim_3feedback %>%
  pivot_wider(names_from="pool", values_from="stock") %>%
  mutate(Feedback_C_Total = C_fast + C_slow +C_passive) 

#Add the one-pool proxy column
sim_3feedback$OnePoolProxy <- sim_Feedback_OnePool$C


#Formatting the data for nice graphing:
sim_3feedback <- sim_3feedback %>% 
                    select(time, Feedback_C_Total, OnePoolProxy) %>%
                    pivot_longer(cols = c("Feedback_C_Total", "OnePoolProxy"),
                                 names_to = "model",
                                 values_to = "stock")


#Plot comparing the two
ggplot(sim_3feedback) +
  geom_line(aes(x=time, y=stock, color=model))


```


###Performance based on IV

```{r Inital Values}

#Creates a simulation using the full feedback model starting at 60% SS
sim_3feedback <- lsoda(y = C_ss_3feedback(parms = century_parameter.ls)*0.6,
                      times = 0:(12*1000),
                      func = dCdt_3feedback,
                      parms = century_parameter.ls) %>%
  as.data.frame() %>%
  pivot_longer(cols = C_fast:C_passive,
               names_to = 'pool',
               values_to = 'stock')%>%
  mutate(pool = factor(pool, levels = c('C_fast', 'C_slow', 'C_passive')),
         model = '3feedback')

#Calculating total C in the Feedback run
sim_3feedback <- sim_3feedback %>%
  pivot_wider(names_from="pool", values_from="stock") %>%
  mutate(Feedback_C_Total = C_fast + C_slow +C_passive)

#Simulation with the one-pool model starting at 60% SS
sim_Feedback_OnePool <- lsoda(y = C_init*0.6,
                times = 0:(12*1000),
                func = dCdt_FeedbackProxy,
                parms = century_parameter.ls) %>%
  as.data.frame()

#Add the one-pool proxy column
sim_3feedback$OnePoolProxy <- sim_Feedback_OnePool$C


#Formatting the data for nice graphing:
sim_3feedback <- sim_3feedback %>% 
                    select(time, Feedback_C_Total, OnePoolProxy) %>%
                    pivot_longer(cols = c("Feedback_C_Total", "OnePoolProxy"),
                                 names_to = "model",
                                 values_to = "stock")


#Plot comparing the two
ggplot(sim_3feedback) +
  geom_line(aes(x=time, y=stock, color=model)) +
  labs(title="Model Comparison from 60% of Steady State", x="Time in months", y="Total Carbon (kg C/(m^2 yr))")
```


```{r Diff_Initial_Values}

Initial_Value_Comparison <- function(partial){
  
  sim_IV <- lsoda(y = C_ss_3feedback(parms = century_parameter.ls)*partial,
                      times = 0:(12*1000),
                      func = dCdt_3feedback,
                      parms = century_parameter.ls) %>%
  as.data.frame() %>%
  mutate(Feedback_C_Total = C_fast + C_slow +C_passive) %>%
  select(time, Feedback_C_Total) 
  

#Simulation with the one-pool model
sim_OP_temp <- lsoda(y = C_init*partial,
                times = 0:(12*1000),
                func = dCdt_FeedbackProxy,
                parms = century_parameter.ls) %>%
  as.data.frame()


#Add the one-pool proxy column
sim_IV$One_Pool_Proxy <- sim_OP_temp$C


Ave_Percent_Diff <- (1/12001)*sum((sim_IV$One_Pool_Proxy-sim_IV$Feedback_C_Total)/sim_IV$Feedback_C_Total)

# ##Attempt at an R squared calculation
# total_error <- sum( ((sum(sim_IV$One_Pool_Proxy)/1001)-sim_IV$Feedback_C_Total)^2)
# residual_error <- sum((sim_IV$One_Pool_Proxy-sim_IV$Feedback_C_Total)^2)
# 
# R_Value <- 1- (residual_error/total_error)

return(Ave_Percent_Diff)
}


Comparisons <- data.frame( Initial_Value_percentage_of_Steady_State= c(0.25,0.5,0.75,0.85,0.875, 0.9375, 0.96875, 0.984373,1),
                            Ave_Percent_Diff = c(Initial_Value_Comparison(partial=0.25),
                                          Initial_Value_Comparison(partial=0.5),
                                          Initial_Value_Comparison(partial=0.75),
                                          Initial_Value_Comparison(partial=0.85),
                                          Initial_Value_Comparison(partial=0.875),
                                          Initial_Value_Comparison(partial=0.9375),
                                          Initial_Value_Comparison(partial=0.96875),
                                          Initial_Value_Comparison(partial=0.984373),
                                          Initial_Value_Comparison(partial=1)
                            )
)

ggplot(Comparisons) + 
    geom_path(aes(x=Initial_Value_percentage_of_Steady_State, y=Ave_Percent_Diff), linetype=3) +   
    geom_point(aes(x=Initial_Value_percentage_of_Steady_State, y=Ave_Percent_Diff), shape=1) +
    labs(title="Model Comparison based on Initial Condition's Distance from Steady State", x="Percentage of Steady State for Initial Condition", y="Average Percent Difference at each time step")






```

### Inputs turned off

How does this aggregate decay rate compare to the full model when soil is removed from its environment, and no new organic carbon enters the system?

```{r InputsOff}
#paramter list with no inputs
no_input_century_parameter.ls <- century_parameter.ls %>% list_assign(inputs=0)

#Creates a simulation using the full feedback model starting at SS, turn off inputs at time 240 (twenty years, monthly time steps)
No_inputs_feedback_sim <- rbind(lsoda(y = C_ss_3feedback(parms = century_parameter.ls),
                                      times = 0:240,
                                      func = dCdt_3feedback,
                                      parms = century_parameter.ls) %>%
                                  as.data.frame(),
                              lsoda(y = C_ss_3feedback(parms = century_parameter.ls),
                                      times = 0:9760,
                                      func = dCdt_3feedback,
                                      parms = no_input_century_parameter.ls) %>%
                                  as.data.frame() %>%
                                  mutate(time=time+240)
                                ) %>%
                          mutate(Feedback_C_Total = C_fast + C_slow +C_passive) %>%
                          mutate(C_ps = C_passive + C_slow) %>%
                          as.data.frame
                


#Simulation with the one-pool model starting at SS, turn off inputs at time 240 (twenty years, monthly time steps)
No_inputs_one_pool_proxy_sim <- rbind(lsoda(y = sum(C_ss_3feedback(parms = century_parameter.ls)),
                                            times = 0:240,
                                            func = dCdt_FeedbackProxy,
                                            parms = century_parameter.ls) %>%
                                       as.data.frame() %>%
                                       set_names(c("time", "One_Pool_Proxy")),
                                      
                                    lsoda(y = sum(C_ss_3feedback(parms = century_parameter.ls)),
                                            times = 0:9760,
                                            func = dCdt_FeedbackProxy,
                                            parms = no_input_century_parameter.ls) %>%
                                      as.data.frame() %>%
                                      set_names(c("time", "One_Pool_Proxy")) %>%
                                      mutate(time=time+240)
                                )
#Makes a dataframe to compare the two simulations total carbon
no_input_model_comparison <- as.data.frame(cbind(No_inputs_feedback_sim$time, No_inputs_feedback_sim$Feedback_C_Total, No_inputs_one_pool_proxy_sim$One_Pool_Proxy)) %>% 
    set_names(c("time", "Feedback_C_Total", "One_Pool_Proxy")) %>%
    pivot_longer(cols = c("Feedback_C_Total", "One_Pool_Proxy"),
                                 names_to = "model",
                                 values_to = "stock")



ggplot(no_input_model_comparison)  +
  # geom_area(data=No_inputs_feedback_sim, aes(x=time, y=C_passive), alpha=0.1, fill="firebrick")+
  # geom_area(data=No_inputs_feedback_sim, aes(x=time, y=C_ps), alpha=0.1, fill="firebrick")+
  # geom_area(data=No_inputs_feedback_sim, aes(x=time, y=Feedback_C_Total), alpha=0.1, fill="firebrick")+
  geom_line(aes(x=time, y=stock, color=model)) +
  geom_vline(xintercept=240, alpha=0.5, linetype="dashed") +
  labs(title="Inputs cut off after 20 years", y="Total Carbon (kg C/(m^2 yr))", x="Time (in months)" )



```
### Fluctuating Inputs

NPP changes on diurnal, seasonal, and even longer-term cycles. When inputs are fluctuating, how does this one-pool proxy respond? We will focus on a yearly seasonal cycle (the parametrizations from Parton are monthly, so this makes sense as a starting point).

```{r InputsFlux}
#Function for simulation with the 3feedback model, adjusting the inputs for seasonal change
dCdt_seasonal_3feedback <- function(t, y, parms){
  pools <- matrix(y, nrow=3)
  allocation_matrix <- matrix(c(parms$input_to_fast, 
                                parms$input_to_slow, 
                                1 - parms$input_to_fast - parms$input_to_slow),
                              nrow = 3)
  decay_matrix <- diag(x= 1/c(parms$turnoverTime_fast,
                             parms$turnoverTime_slow,
                             parms$turnoverTime_passive))
  transfer_matrix <- matrix(c(1,  -parms$slow_to_fast, -parms$passive_to_fast,
                              -parms$fast_to_slow, 1, -parms$passive_to_slow,
                              -parms$fast_to_passive, -parms$slow_to_passive,  1),
                            nrow = 3,
                            byrow=TRUE)
  #dcdt = inputs * allocation + transfer_matrix %*% decay_matrix %*% y
  #decay_matrix %*% transfer_matrix %*% y = - inputs * allocation
  
  flux_input <- (5/300)*sin((pi/6)*t+(3*pi/2))+(5/300)
  
  ans <- flux_input * allocation_matrix - decay_matrix %*% transfer_matrix %*% y
  
  return(list(c(C_fast = ans[1],
           C_slow = ans[2],
           C_passive = ans[3])))
}


#Demonstration simulation that the method works to create seasonal change, over 5 years
seasonal_sim_3feedback <- lsoda(y = C_ss_3feedback(parms = century_parameter.ls),
                                      times = 0:60,
                                      func = dCdt_seasonal_3feedback,
                                      parms = century_parameter.ls) %>%
                                  as.data.frame() %>% 
                                  mutate(Feedback_C_total = C_fast+C_slow+C_passive)

                                  

ggplot(pivot_longer(data=seasonal_sim_3feedback, cols = C_fast:C_passive, names_to = "pool", values_to = "stock")) +
  geom_line(aes(x=time, y=stock, color=pool)) +
  geom_line(aes(x=time, y=Feedback_C_total))


#One-Pool rate change function, with seasonal rate of change
  ##The calucation of the Proxy decay rate relies on a steady state solution, which technically does not exist when inputs change - we then assume that the average over the year is constant, and that the steady state solution suffices.


dCdt_seasonal_FeedbackProxy <- function(t, y, parms){
  flux_input <- (5/300)*sin((pi/6)*t+(3*pi/2))+(5/300)
  ans <- flux_input - y / parms$proxy_decay_rate
  return(list(ans))
}


#Demonstration simulation that the method works to create seasonal change, over 5 years
seasonal_sim_FeedbackProxy <- lsoda(y = C_init,
                                      times = 0:60,
                                      func = dCdt_seasonal_FeedbackProxy,
                                      parms = century_parameter.ls) %>%
                                  as.data.frame() %>%
                                  rename(One_Pool_Proxy = C)

ggplot(seasonal_sim_FeedbackProxy) +
  geom_line(aes(x=time, y=One_Pool_Proxy))

                                
#Makes a dataframe to compare the two simulations total carbon
seasonal_inputs_model_comparison <- as.data.frame(cbind(seasonal_sim_3feedback$time, seasonal_sim_3feedback$Feedback_C_total, seasonal_sim_FeedbackProxy$One_Pool_Proxy)) %>% 
    set_names(c("time", "Feedback_C_Total", "One_Pool_Proxy")) %>%
    pivot_longer(cols = c("Feedback_C_Total", "One_Pool_Proxy"),
                                 names_to = "model",
                                 values_to = "stock")



ggplot(seasonal_inputs_model_comparison)  +
  geom_line(aes(x=time, y=stock, color=model)) +
  labs(title="Seasonal Variation over 5 Years", y="Total Carbon (kg C/(m^2))", x="Time (in months)" )



```
Holy crap that's close!!!!

```{r seasonal_comparison_stats}

#How different are they?
seasonal_Ave_Percent_Diff <- (1/61)*sum((seasonal_sim_FeedbackProxy$One_Pool_Proxy-seasonal_sim_3feedback$Feedback_C_Total)/seasonal_sim_3feedback$Feedback_C_Total)

print(seasonal_Ave_Percent_Diff)
print("That's 5 thousandths of a percent average difference, or, the One Pool Proxy decay rate undershoots the amount of carbon .005% of the Full Feedback model on average.")

```


###Increasing Inputs

