---
title: "Regime Shift Scenarios"
author: "Ashley Bonner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, echo=FALSE}

library(tidyverse) #data library
library(deSolve) #ODE solver
library(patchwork) #figure arrangement
library(dplyr) #data wrangling

knitr::opts_chunk$set(echo = TRUE)

```


# Scenario Goals
We are looking to compare a one-pool model with a single decay rate against a multi-pool model with transfer of soil carbon between pools in simulated real-world scenarios.
This Markdown is going to test how a one-pool model fares against a multipool soil carbon model to a vegetation regime shift

First, setting up our models for testing.

```{r setup}
#functions
source("R/AggregateDecayRate.R")
source("R/OnePool_ODE.R")
source("R/FivePool_ODE.R")
source("R/FivePool_SS.R")

#parameters
##five pool
temp <- read_csv(file.path("Data/FivePoolModel_Parameters.csv"), show_col_types = FALSE)
Params_FivePool.ls <- list()

for(i in 1:ncol(temp)) {
  Params_FivePool.ls[[i]] <- as.numeric(temp[,i])
}
names(Params_FivePool.ls) <- colnames(temp)

#create the One-pool parameters with aggregate decay rate.
Params_Aggregate.ls <- list(ave_inputs = Params_FivePool.ls$ave_inputs, #g-C m-2 yr-1, according to CENTURY model validation
                            turnoverTime = Proxy_decayrate.fn(parms = Params_FivePool.ls)
                            )

```



# Regime Shift - new input levels

Modeling soil carbon in earth systems (such as forecasting carbon totals for climate change) want to capture how soil carbon stocks respond to natural and unnatural dynamics and fluctuations.
In temperate systems, a seasonal fluctuation in Vegetation Biomass' Net Primary Productivity is well established.
As carbon inputs to soil are assumed to follow NPP fluctuations, we want to assess how our one pool model responds to these fluctuations, to determine if a one pool model can suffice in global earth system models.
We'll use a weekly simulation over five years to determine fit.


```{r InputsForIncub}
##define a function where inputs shift after a year
NPPshift.fn <- function(time, parms){
  
  if(time < 0.25){
    u <- parms$ave_inputs
  }else{
    u <- parms$ave_inputs*parms$NPPshift
  }
  
  return(u)
}

# #Simulation Times vector
SimTimes <- c(seq(0, 1, by = 1/12),
              seq(1, 10, by = 1),
              seq(10, 100, by = 10),
              seq(100, 1e4, by = 100),
              seq(1e4, 1e6, by = 1e4)) %>%
  unique()
  
  # seq(0, 1e6, by = 1)

Inputs_sim <- as.data.frame(SimTimes)

NPP_shifts <- c(0.5, 0.9, 1, 1.1, 1.5)

for(i in 1:length(NPP_shifts)){

Params_Aggregate.ls$NPPshift <- NPP_shifts[i] 

temp <- as.data.frame(SimTimes) %>%
  mutate(inputs = NPPshift.fn(time = SimTimes, parms = Params_Aggregate.ls), 
         .by=SimTimes) %>%
  mutate(inputs_shift = as.character(NPP_shifts[i]) )

Inputs_sim <- full_join(Inputs_sim, temp)
}

ggplot(Inputs_sim) +
  geom_line(aes(x = SimTimes, y=inputs, color=inputs_shift)) +
  labs(title = "Inputs over Simulation period", x = "Time (years)", y = "Soil Carbon Inputs (gC per m2 per year)")
```


```{r Simulations}
###AGGREGATE DECAY RATE ONE POOL SIMULATIONS
##add needed input function to parameter lists
Params_Aggregate.ls$inputs.fn <- NPPshift.fn

#creating an empty data frame to populate through for loop
NPPshift_Sim1.df <- data.frame(time=as.numeric(),
                               Soil_Carbon_Concentration=as.numeric(),
                               Respiration_Rate=as.numeric(),
                               Input_Rate=as.numeric(),
                               inputs_shift=as.character())

for(i in 1:length(NPP_shifts)){

Params_Aggregate.ls$NPPshift <- NPP_shifts[i] 

 temp <- lsoda(y = c("Cumulative_Respiration" = 0, 
                                  "Soil_Carbon_Concentration" = sum(FivePool_SS.fn(parms = Params_FivePool.ls))),
                times = SimTimes,
                func = OnePool_ODE.fn,
                parms = Params_Aggregate.ls) %>%
  as.data.frame() %>%
  select(!c(Cumulative_Respiration)) %>%
  mutate(Respiration_Rate = OnePool_ODE.fn(t = 1, 
                                           y = c(NA, Soil_Carbon_Concentration),
                                           parms = Params_Aggregate.ls
                            )[[1]]['Cumulative_Respiration'],
          .by = everything()
    ) %>%
  mutate(Input_Rate = Params_Aggregate.ls$inputs.fn(t = time,
                                                  parms = Params_Aggregate.ls),
         .by = everything()) %>%
  mutate(inputs_shift = as.character(NPP_shifts[i]) )
 
 NPPshift_Sim1.df <- full_join(NPPshift_Sim1.df, 
                               temp, 
                               by = join_by(time, Soil_Carbon_Concentration, Respiration_Rate, Input_Rate, inputs_shift))

}

##Graph both shifts with the one pool model
ggplot(NPPshift_Sim1.df %>% pivot_longer(cols = Soil_Carbon_Concentration:Input_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass, color=inputs_shift)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  labs(title = "Seasonal Simulation: One Pool", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)")


####FIVE POOL SIMULATIONS
##Add needed functino to parameter list for Five Pool simulations
Params_FivePool.ls$inputs.fn <- NPPshift.fn


#creating an empty data frame to populate through for loop
NPPshift_Sim5.df <- data.frame(time=as.numeric(),
                               Metabolic_Plant_Residue=as.numeric(),
                               Structural_Plant_Residue=as.numeric(),
                               Fast_Soil=as.numeric(),
                               Slow_Soil=as.numeric(),
                               Passive_Soil=as.numeric(),
                               Soil_Carbon_Concentration=as.numeric(),
                               Respiration_Rate=as.numeric(),
                               Input_Rate=as.numeric(),
                               inputs_shift=as.character())

for(i in 1:length(NPP_shifts)){

Params_FivePool.ls$NPPshift <- NPP_shifts[i] 

 temp <- lsoda(y = c("Cumulative_Respiration" = 0, 
                                  FivePool_SS.fn(parms = Params_FivePool.ls)),
                times = SimTimes,
                func = FivePool_ODE.fn,
                parms = Params_FivePool.ls) %>%
  as.data.frame() %>%
  select(!c(Cumulative_Respiration)) %>%
  mutate(Soil_Carbon_Concentration = Metabolic_Plant_Residue + Structural_Plant_Residue + Fast_Soil + Slow_Soil + Passive_Soil) %>%
  mutate(Respiration_Rate = FivePool_ODE.fn(t = 1, 
                                           y = c("Cumulative_Respiration" = 0, 
                                                Metabolic_Plant_Residue = Metabolic_Plant_Residue,
                                                Structural_Plant_Residue = Structural_Plant_Residue,
                                                Fast_Soil = Fast_Soil,
                                                Slow_Soil = Slow_Soil,
                                                Passive_Soil = Passive_Soil),
                                            parms = Params_FivePool.ls
                                  )[[1]]['Cumulative_Respiration'],
          .by = everything()
    )%>%
  mutate(Input_Rate = Params_FivePool.ls$inputs.fn(t = time,
                                                  parms = Params_FivePool.ls),
         .by = everything()
  ) %>%
  mutate(inputs_shift = as.character(NPP_shifts[i]) )
 
  NPPshift_Sim5.df <- full_join(NPPshift_Sim5.df, 
                               temp,
                               by = join_by(time, Metabolic_Plant_Residue, Structural_Plant_Residue, Fast_Soil, Slow_Soil, Passive_Soil, Soil_Carbon_Concentration, Respiration_Rate, Input_Rate, inputs_shift))
 
}

ggplot(NPPshift_Sim5.df %>% pivot_longer(cols = Metabolic_Plant_Residue:Input_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass, color = inputs_shift)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  labs(title = "Seasonal Simulation:Five Pools", subtitle = "Carbon Stocks by Pool", x = "Time (years)", y = "SOC (gC per m2) or Rate  (gC per m2 per year)")


ggplot(NPPshift_Sim5.df %>%
         mutate(Plant_Residue = Metabolic_Plant_Residue + Structural_Plant_Residue) %>%
         select(!c('Respiration_Rate', 'Metabolic_Plant_Residue', 'Structural_Plant_Residue')) %>%
         pivot_longer(cols = c('Fast_Soil', 'Slow_Soil', 'Passive_Soil', 'Plant_Residue', 'Soil_Carbon_Concentration'),
                                  names_to = 'Pool',
                                  values_to = 'Mass')) +
  geom_line(aes(x=time, y=Mass, color = Pool, linetype = inputs_shift)) + 
  labs(title = "Seasonal Simulation: Five Pools", subtitle = "Carbon Stocks by Pool", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)")



ggplot(NPPshift_Sim1.df) + 
  geom_line(aes(x=time, 
                y=Soil_Carbon_Concentration/sum(FivePool_SS.fn(parms = Params_FivePool.ls)),
                color=inputs_shift), linetype="dashed") +
  geom_line(data = NPPshift_Sim5.df, aes(x=time, y=Soil_Carbon_Concentration/sum(FivePool_SS.fn(parms = Params_FivePool.ls)), color=inputs_shift), linetype="dotted") + 
  scale_x_log10()

temp <- NPPshift_Sim5.df%>% 
         select(time, Metabolic_Plant_Residue, Structural_Plant_Residue, Fast_Soil, Slow_Soil, Passive_Soil, inputs_shift) %>%
         pivot_longer(cols = Metabolic_Plant_Residue:Passive_Soil,
                                  names_to = 'Pool',
                                  values_to = 'Mass')

temp$Pool <- factor(temp$Pool, levels = c("Metabolic_Plant_Residue", "Structural_Plant_Residue", "Fast_Soil", "Slow_Soil", "Passive_Soil"))

ggplot(temp) + 
  geom_line(aes(x=time, y=Mass, color = inputs_shift)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') 

ggplot(temp) +
  geom_area(aes(x=time, y=Mass, fill = Pool)) +
  scale_fill_brewer(palette = "Spectral") +
  facet_wrap(~ inputs_shift, ncol=1, scales='free') 


```


```{r}



```






#### BELOW NEEDS UPDATING  - MIGHT GET CUT ALTOGETHER

# Model to Model Comparisons

```{r}
NPPshift_Sim_comp.df <- rbind(NPPshift_Sim1.df %>% 
            pivot_longer(cols = c("Respiration_Rate", "Input_Rate"),
                         names_to = "Rate_Type",
                         values_to = "Rate") %>%
            mutate(Model = case_when(Rate_Type == "Input_Rate" ~ "Input",
                                     Rate_Type == "Respiration_Rate" ~ "OnePool")) %>%
            select(c("time", "Soil_Carbon_Concentration", "Rate", "Model", "Shift")),
           NPPshift_Sim5.df %>% 
             rename(Rate = Respiration_Rate) %>%
             select(c("time", "Rate", "Soil_Carbon_Concentration", "Shift")) %>% 
             mutate(Model = "FivePool")
           )

ggplot(NPPshift_Sim_comp.df) +
  geom_line(aes(x=time, y=Soil_Carbon_Concentration, color=Model, linetype = Shift)) +
  labs(title = "Seasonal Simulation", subtitle = "Carbon Stocks", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)")

ggplot(NPPshift_Sim_comp.df ) +
  geom_line(aes(x=time, y=Rate, color=Model, linetype = Shift)) +
  labs(title = "Seasonal Simulation", subtitle = "Respiration", x = "Time (years)", y = "Respiration Rate (gC per m2 per year)")


ggplot(NPPshift_Sim_comp.df %>% 
        select(c(time, Model, Soil_Carbon_Concentration, Shift))%>%
         pivot_wider(names_from = Model,
                     values_from = Soil_Carbon_Concentration)
       ) +
  geom_point(aes(x=FivePool, y = OnePool, color = Shift)) +
  geom_abline(color='gray') 

ggplot(NPPshift_Sim_comp.df %>% 
        select(c(time, Model, Rate, Shift)) %>%
         pivot_wider(names_from = Model,
                     values_from = Rate) %>%
         select(!Input)
       ) +
  geom_point(aes(x=FivePool, y = OnePool, color = Shift)) +
  geom_abline(color='gray') 
```


