---
title: "Seasonal Scenario"
author: "Ashley Bonner"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse) #data library
library(deSolve) #ODE solver
library(patchwork) #figure arrangement
library(dplyr) #data wrangling

knitr::opts_chunk$set(echo = TRUE)
```


# Scenario Goals
We are looking to compare a one-pool model with a single decay rate against a mulit-pool model with transfer of soil carbon between pools in simulated real-world scenarios.
This Markdown is going to test our method in a simulated seasonal fluctuation scenario.

First, setting up our models for testing.

```{r setup}
#functions
source("R/AggregateDecayRate.R")
source("R/OnePool_ODE.R")
source("R/FivePool_ODE.R")
source("R/FivePool_SS.R")

#parameters
##five pool

# light_sens_figs.df <- read_csv(file.path(
#   "DeAD_super_simp_light_sens_20240918.csv")) %>%
#   merge(parms_light_sens2, by = "index")


Params_FivePool.ls <- list(
                  ##Input-based parameters (such as allocation)
                  ave_inputs = 260,
                        #CENUTRY model uses 260 grams C per m^2 per year
                        #u; kg-carbon m-2 yr-1, 
                    input_to_struc = 2/9,  
                    input_to_meta = 7/9,  
                        #C goes through litter pools first
                    input_to_fast = 0,  
                    input_to_slow = 0,  
                    input_to_passive = 0,  
                            
                  ##turnover times (years)
                     turnoverTime_meta = 3.5,  #this is CENTURY 1987, CENTURY 1983 has .1 to 1 year
                     turnoverTime_struc = 9.5, #adjusted out of initial range to match CENTURY model validation for total plant residue
                        # CENTURY: 1 to 5 years
                     turnoverTime_fast = 10, #this is CENTURY 1987, CENTURY 1983 has 10 years
                     turnoverTime_slow = 50, #adjusted, CENTURY 1987 has ~263, CENTURY 1983 has 50
                     turnoverTime_passive = 2500, #adjusted, CENTURY 1987 has ~7690, CENTURY 1983 has 1500 years
                     
                  ##Transfer Matrix values 
                       struc_to_fast = .7*.35, 
                       struc_to_slow = .3*.35, 
                          #60% lost to respiration, the remaining 40% is called Net Microbial Production 
                          #Adjusted to 75% lost to respiration to reach model validations percentages
                       meta_to_fast = .35,  
                          #60% lost to respiration, rest to active
                          #Adjusted to 75% lost to respiration to reach model validations percentages
                       fast_to_slow = .4,      
                          #60% lost to respiration, rest to slow
                       slow_to_fast = 0.92*.4,        
                       slow_to_passive = .08*.4,     
                          #60% lost to respiration, rest split between active and passive
                       passive_to_fast = .4,
                    #.4,
                          #60% lost to respiration, rest to slow
                    
                      #all the zero-flows
                        meta_to_struc = 0,
                        meta_to_slow = 0,
                        meta_to_passive = 0,
                        struc_to_meta = 0,
                        struc_to_passive = 0,
                        fast_to_passive = 0,
                        fast_to_meta = 0,
                        slow_to_meta = 0,
                        passive_to_meta = 0,
                        fast_to_struc = 0,
                        slow_to_struc = 0,     
                        passive_to_slow = 0,
                        passive_to_struc = 0
                       )

#create the One-pool parameters with aggregate decay rate.
Params_Aggregate.ls <- list(ave_inputs = 260, #g-C m-2 yr-1, according to CENTURY model validation
                            turnoverTime = Proxy_decayrate.fn(parms = Params_FivePool.ls)
                            )

```

# Seasonal Variance

Modeling soil carbon in earth systems (such as forecasting carbon totals for climate change) want to capture how soil carbon stocks respond to natural and unnatural dynamics and fluctuations.
In temperate systems, a seasonal fluctuation in Vegetation Biomass' Net Primary Productivity is well established.
As carbon inputs to soil are assumed to follow NPP fluctuations, we want to assess how our one pool model responds to these fluctuations, to determine if a one pool model can suffice in global earth system models.
We'll use a weekly simulation over five years to determine fit.


```{r InputsForIncub}
##define a no-inputs function 
seasonal.fn <- function(time, parms){
  Amplitude <- parms$ave_inputs
  Horizontal_Shift <- 0.25 #0 #0.5 #simulation starts
  Period <- 1 #Seasons are yearly
    B <- 2*pi*Period
    C <- Horizontal_Shift*B
  Vertical_Shift <- parms$ave_inputs #Shifts function up so lowest NPP is zero.
  
  u <- Amplitude*sin(B*time+C) + Vertical_Shift 
  
  return(u)
}

#Simulation Times vector
SimTimes <- seq(0, 5, by = 1/104) 

Inputs_sim <- as.data.frame(SimTimes)%>%
  mutate(inputs = seasonal.fn(time = SimTimes, parms = Params_Aggregate.ls))

Inputs.gr <- ggplot(Inputs_sim) + 
  geom_line(aes(x = SimTimes, y=inputs), size = 1.5) + 
  labs(title = "Inputs over Simulation period", x = "Time (years)", y = "Soil Carbon Inputs (gC per m2 per year)")

Inputs.gr +
  labs(title = "Seasonal NPP") + 
  theme_bw() +  
theme(plot.title = element_text(size = 20),  
strip.text = element_text(size=18),  
axis.text = element_text(size=15),  
axis.title = element_text(size=18),  
legend.position="bottom",  
legend.text=element_text(size=15),  
legend.title=element_text(size=18))  

ggsave("SeasonalInputs.pdf", height=5, width=12)

```


```{r Simulations}
##add needed input function to parameter lists
Params_Aggregate.ls$inputs.fn <- seasonal.fn
Params_FivePool.ls$inputs.fn <- seasonal.fn

Seasonal_Sim1.df <- lsoda(y = c("Cumulative_Respiration" = 0, 
                                  "Soil_Carbon_Concentration" = sum(FivePool_SS.fn(parms = Params_FivePool.ls))),
                times = SimTimes,
                func = OnePool_ODE.fn,
                parms = Params_Aggregate.ls) %>%
  as.data.frame() %>%
  select(!c(Cumulative_Respiration)) %>%
  mutate(Respiration_Rate = OnePool_ODE.fn(t = NA, 
                                           y = c(NA, Soil_Carbon_Concentration),
                                           parms = Params_Aggregate.ls
                            )[[1]]['Cumulative_Respiration'],
          .by = everything()
    )%>%
  mutate(Input_Rate = Params_Aggregate.ls$inputs.fn(t = time,
                                                  parms = Params_Aggregate.ls),
         .by = everything()
  )

ggplot(Seasonal_Sim1.df %>% pivot_longer(cols = Soil_Carbon_Concentration:Input_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  labs(title = "Seasonal Simulation: One Pool", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)")



Seasonal_Sim5.df <- lsoda(y = c("Cumulative_Respiration" = 0, 
                                  FivePool_SS.fn(parms = Params_FivePool.ls)),
                times = SimTimes,
                func = FivePool_ODE.fn,
                parms = Params_FivePool.ls) %>%
  as.data.frame() %>%
  select(!c(Cumulative_Respiration)) %>%
  mutate(Soil_Carbon_Concentration = Metabolic_Plant_Residue + Structural_Plant_Residue + Fast_Soil + Slow_Soil + Passive_Soil) %>%
  mutate(Respiration_Rate = FivePool_ODE.fn(t = time, 
                                           y = c("Cumulative_Respiration" = 0, 
                                                Metabolic_Plant_Residue = Metabolic_Plant_Residue,
                                                Structural_Plant_Residue = Structural_Plant_Residue,
                                                Fast_Soil = Fast_Soil,
                                                Slow_Soil = Slow_Soil,
                                                Passive_Soil = Passive_Soil),
                                            parms = Params_FivePool.ls
                                  )[[1]]['Cumulative_Respiration'],
          .by = everything()
    )%>%
  mutate(Input_Rate = Params_FivePool.ls$inputs.fn(t = time,
                                                  parms = Params_FivePool.ls),
         .by = everything()
  )

ggplot(Seasonal_Sim5.df %>% pivot_longer(cols = Metabolic_Plant_Residue:Input_Rate,
                               names_to = "Pool",
                               values_to = "Mass")) + 
  geom_line(aes(x=time, y=Mass)) + 
  facet_wrap(~ Pool, ncol=1, scales='free') + 
  labs(title = "Seasonal Simulation:Five Pools", subtitle = "Carbon Stocks by Pool", x = "Time (years)", y = "SOC (gC per m2) or Rate  (gC per m2 per year)")


ggplot(Seasonal_Sim5.df %>%
         mutate(Plant_Residue = Metabolic_Plant_Residue + Structural_Plant_Residue) %>%
         select(!c('Respiration_Rate', 'Metabolic_Plant_Residue', 'Structural_Plant_Residue')) %>%
         pivot_longer(cols = Fast_Soil:Plant_Residue,
                                  names_to = 'Pool',
                                  values_to = 'Mass')) +
  geom_line(aes(x=time, y=Mass, linetype = Pool, color = Pool)) + 
  labs(title = "Seasonal Simulation: Five Pools", subtitle = "Carbon Stocks by Pool", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)")


```


# Model to Model Comparisons

```{r}
Seasonal_Sim_comp.df <- rbind(Seasonal_Sim1.df %>% 
            pivot_longer(cols = c("Respiration_Rate", "Input_Rate"),
                         names_to = "Rate_Type",
                         values_to = "Rate") %>%
            mutate(Model = case_when(Rate_Type == "Input_Rate" ~ "Input",
                                     Rate_Type == "Respiration_Rate" ~ "OnePool")) %>%
            select(c("time", "Soil_Carbon_Concentration", "Rate", "Model")),
           Seasonal_Sim5.df %>% 
             rename(Rate = Respiration_Rate) %>%
             select(c("time", "Rate", "Soil_Carbon_Concentration")) %>% 
             mutate(Model = "FivePool")
           )

Seas_Stocks_time.plot <- ggplot(Seasonal_Sim_comp.df %>%
                                  filter(Model != "Input")) +
  geom_line(aes(x=time, y=Soil_Carbon_Concentration, color=Model, linetype = Model), size = 2, alpha = 0.9) + 
  scale_color_manual(values = c("orange", "navyblue"))+
  labs(x = "Time (years)", y = "SOC (gC per m2)")

Seas_Stocks_time.plot +
  labs(title = "Seasonal Simulation") + 
  theme_bw() +  
theme(plot.title = element_text(size = 20),  
strip.text = element_text(size=18),  
axis.text = element_text(size=15),  
axis.title = element_text(size=18),  
legend.position="bottom",  
legend.text=element_text(size=15),  
legend.title=element_text(size=18))  

ggsave("Seasonal.pdf", height=9, width=12)

ggplot(Seasonal_Sim_comp.df %>% filter(Model != "Input")) +
  geom_line(aes(x=time, y=Soil_Carbon_Concentration, color=Model)) +
  labs(title = "Seasonal Simulation", subtitle = "Carbon Stocks by model", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)") +
  facet_wrap(~Model, ncol=1, scales = 'free')

Seas_Resp_time.plot <- ggplot(Seasonal_Sim_comp.df ) +
  geom_line(aes(x=time, y=Rate, color=Model), size = 1.5) +
  scale_color_manual(values = c("orange", "green3", "navyblue"))+
  labs(x = "Time (years)", y = "Respiration Rate (gC per m2 per year)")

Seas_Resp_time.plot +
  labs(title = "Seasonal Simulation") + 
  theme_bw() +  
theme(plot.title = element_text(size = 20),  
strip.text = element_text(size=18),  
axis.text = element_text(size=15),  
axis.title = element_text(size=18),  
legend.position="bottom",  
legend.text=element_text(size=15),  
legend.title=element_text(size=18))  

ggsave("SeasonalResp.pdf", height=9, width=12)

Seas_Resp_time.plot +
  labs(title = "Seasonal Simulation", subtitle = "Respiration")

ggplot(Seasonal_Sim_comp.df ) +
  geom_line(aes(x=time, y=Rate, color=Model)) +
  labs(title = "Seasonal Simulation", subtitle = "Rates (Respiration & Input rate)", x = "Time (years)", y = "Respiration Rate (gC per m2 per year)")+
  facet_wrap(~Model, ncol=1, scales = 'free')


#Comparison graphs, using a temp dataframe 

temp <- Seasonal_Sim_comp.df %>% 
        select(c(time, Model, Soil_Carbon_Concentration))%>%
         pivot_wider(names_from = Model,
                     values_from = Soil_Carbon_Concentration) %>%
        select(-"Input") %>%
  mutate("Difference" = OnePool - FivePool) %>%
  mutate("DifferenceRelativeToFivePoolModel" = (Difference/FivePool)*100) 

seasonal_low <- min(temp["OnePool"], temp["FivePool"])
seasonal_high <- max(temp["OnePool"], temp["FivePool"])

Seas_Stocks_M2M.plot <- ggplot(temp %>%
  filter(time <= 1)) +
  geom_point(aes(x=FivePool, y = OnePool), alpha=0.5) +
  geom_abline(color='red') +
  lims(x=c(seasonal_low,seasonal_high), y=c(seasonal_low,seasonal_high)) + 
  labs(title = 'Seasonal Variation SOC stock', subtitle = "(gC per m2)")

SeasonPercentage.df <- data.frame(temp["time"], temp["DifferenceRelativeToFivePoolModel"])

temp <- Seasonal_Sim_comp.df %>% 
        select(c(time, Model, Rate))%>%
         pivot_wider(names_from = Model,
                     values_from = Rate) %>%
  mutate("Difference" = OnePool - FivePool) %>%
  mutate("DifferenceRelativeToInputRate" = (Difference/FivePool)*100) 

seasonal_low <- min(temp["OnePool"], temp["FivePool"])
seasonal_high <- max(temp["OnePool"], temp["FivePool"])

Seas_Resp_M2M.plot <- ggplot(temp %>%
  filter(time <= 1)) +
  geom_point(aes(x=FivePool, y = OnePool), alpha=0.5) +
  geom_abline(color='red') +
  lims(x=c(seasonal_low,seasonal_high), y=c(seasonal_low,seasonal_high)) + 
  labs(title = 'Seasonal Variation Respiration Rate', subtitle = "(gC per m2 per year)")

Seas_Stocks_M2M.plot
Seas_Resp_M2M.plot

ggplot(temp)+
  geom_line(aes(x=time, y=DifferenceRelativeToInputRate))


ggplot(IncubPercentage.df %>% filter(time <=1) ) + 
  geom_line(aes(x=time, y=DifferenceRelativeToFivePoolModel)) + 
  geom_line(data = SeasonPercentage.df %>% filter(time <=1), aes(x=time, y=DifferenceRelativeToFivePoolModel)) 
```


#Breaking out soil and litter

```{r eval=FALSE}
Seasonal_Sim5_Breakdown <- Seasonal_Sim5.df %>% 
  mutate(SOC = Fast_Soil + Slow_Soil + Passive_Soil,
         Plant_Residue = Metabolic_Plant_Residue + Structural_Plant_Residue) %>%
  select(c("time", "Plant_Residue", "SOC"))

ggplot(Seasonal_Sim5_Breakdown %>%
         pivot_longer(cols = c("Plant_Residue", 
                               "SOC"),
                      names_to = "Carbon_Type",
                      values_to = "Concentration")) +
  geom_line(aes(x=time, y=Concentration))+
  facet_wrap(~ Carbon_Type, ncol=1) 

ggplot(Seasonal_Sim5_Breakdown %>%
         pivot_longer(cols = c("Plant_Residue", 
                               "SOC"),
                      names_to = "Carbon_Type",
                      values_to = "Concentration")) +
  geom_area(aes(x=time, y=Concentration, fill = Carbon_Type)) 

Seasonal_Sim_comp2.df <- rbind(Seasonal_Sim1.df %>% 
              mutate(Model ="OnePool") %>%
              rename("SOC" = "Soil_Carbon_Concentration") %>%
              select(c("time", "SOC", "Model")),
           Seasonal_Sim5_Breakdown %>%
             select(c("time", "SOC")) %>% 
             mutate(Model = "FivePool")
           )

ggplot(Seasonal_Sim_comp2.df) +
  geom_line(aes(x=time, y=SOC, color=Model)) +
  labs(title = "Seasonal Simulation", subtitle = "Carbon Stocks", x = "Time (years)", y = "Soil Carbon Concentration (gC per m2)")

```
